\section{Introduction}
    This chapter introduces the Bellman optimal equation (BOE), and proves serveral
    properties of it. This section mainly follows the book \cite{Zhao2025}.
\section{Optimal Policy}
    The reinforcement learning algorithms are all designed to find an optimal policy for
    given enviroment. Then, what is an 'optimal' policy? The definition is given below.
    \begin{defn}{Optimal Policy and Optimal State Value}{optPol}
        A policy $\pi^*$ is optimal if and only if $v_\pi^*(s) \geq v_\pi(s)$ for all $s \in \mathcal{S}$
        and for any other policy $\pi$. The state values of $\pi^*$ are the optimal state values.
    \end{defn}
    \begin{defn}{Better Policy}{betterPol}
        A policy $\pi_1$ is \textbf{better} than a policy $\pi_2$ if and only if
        $v_{\pi_1}(s) \geq v_{\pi_2}(s)$ for all $s \in \mathcal{S}$.
    \end{defn}
    The questions that naturally arise are:
    \begin{itemize}
        \item Existence: Does the optimal policy exists?
        \item Uniqueness: Is the optimal policy unique?
        \item Stochasticity: Is the optimal policy stochastic or deterministic?
        \item Algorithm: How do we obtain the optimal policy and the optimal state values?
    \end{itemize}
\section{Bellman Optimality Equation}
    The optimal policy and optimal state values can be analyised via the \textbf{Bellman optimality equation} (BOE).
    The optimal state values satisfies following Bellman optimality equation. This claim in fact has two logically independent statements:
    \begin{enumerate}
        \item The BOE, viewed purely as a fixed-point equation, has a unique solution (Section \ref{sec:SolveBellmanOptEq});
        \item The unique solution of BOE concides with the optimal stat-value function $v^*$, and the derived greedy policy is optimal (Section \ref{sec:OptPolOptStateValBOE})
    \end{enumerate}
    Note that the Section \ref{sec:SolveBellmanOptEq} does not assume that the state-value function is optimal.
    \begin{defn}{Bellman Optimality Equation}{BellOptEq}
        Let $\Pi$ be a set of possible policies and let $\gamma \in (0,1)$ be a discount rate. Then, the Bellman Optimality Equation is defined as
        \begin{align}
            v(s) &= \max_{\pi \in \Pi} \sum_{a \in \mathcal{A}(s)}\pi(a|s)\left( \sum_{r \in \mathcal{R}}p(r|s,a)r + \gamma \sum_{s' \in \mathcal{S}}p(s'|s,a)v(s') \right) \\
                 &= \max_{\pi \in \Pi} \sum_{a \in \mathcal{A}(s)}\pi(a|s)q(s,a)
        \end{align}
        where $v(s),v(s')$ are unkown variables to be solved.
        Additionally, the matrix-vector form of Bellman optimality equation is given as
        \begin{equation}
            v = \max_{\pi \in \Pi}(r_\pi + \gamma P_\pi v)
        \end{equation}
        where $r_\pi$ and $P_\pi$ are given as \ref{eq:VecForm}.
    \end{defn}
    Using the matrix-vector form of BOE, we can express BOE very simply as following:
    \begin{equation}
        v = f(v)
    \end{equation}
    where $f(v) = \max_{\pi \in \Pi}(r_\pi + \gamma P_\pi v)$ \label{eq:BOERightHandSide}.
\section{Solving The Bellman Optimality Equation}\label{sec:SolveBellmanOptEq}
    \subsection{Contraction Mapping Theorem}
        We propose a \textbf{contraction mapping theorem}, the main tool for solving the BOE.
        The contraction mapping theorem is a powerful tool for analyzing general nonlinear equations.
        \begin{defn}{Contraction Mapping}{ContractMap}
            Let $f : \mathbb{R}^d \to \mathbb{R}^d$. Then, the function $f$ is a \textbf{contraction mapping}
            if there exists $\gamma \in (0,1)$ such that
            \begin{equation}
                \|f(x_1) - f(x_2)\| \leq \gamma \|x_1-x_2\|
            \end{equation}
            for all $x_1,x_2 \in \mathbb{R}^d$.
        \end{defn}
        Now we present the contraction mapping theorem.
        \begin{thm}{Contraction Mapping Theorem}{ContractMap}
            Let $f:\mathbb{R}^n \to \mathbb{R}^n$ be a contraction mapping. Then, the following properties hold for equation $x=f(x)$:
            \begin{itemize}
                \item Existence: There exists a fixed point $x^*$ satisfying $f(x^*)=x^*$;
                \item Uniqueness: The fixed point $x^*$ is unique;
                \item Algorithm: The following sequence $\{x_k\}$ converges to the fixed point $x^*$ exponentially, for any initial $x_0$.
                    \begin{equation}
                        x_{k+1} = f(x_k)
                    \end{equation}
            \end{itemize}
        \end{thm}
        \begin{proof}
            \textit{Part 1.} (Convergence) The sequence $\{x_k\}$ defined as $x_{k+1}=f(x_k)$ converges.

            We show that the sequence $\{x_k\}$ is a Cauchy sequence; i.e., there always exists an $N \in \mathcal{N}$ such that $\|x_m-x_n\| < \epsilon$ for any $\epsilon > 0$.
            The Cauchy sequence has a nice property; the Cauchy sequence always converges. This can be proved by using the properties of $\mathcal{R}$, that the $\mathcal{R}$ is a complete set.
            We won't show this here. The following holds:
            \begin{align}
                \|x_{k+1}-x_k\| = &\|f(x_k) - f(x_{k-1})\| \\
                                  &\leq \gamma \|x_k - x_{k-1}\| \\
                                  &\leq \gamma^2 \|x_{k-1}-x_{k-2}\| \\
                                  &\leq \gamma^3 \|x_{k-2}-x_{k-3}\| \\
                                  &\vdots \\
                                  &\leq \gamma^k \|x_1 - x_0\|
            \end{align}
            Using this, we can show that
            \begin{align}
                \|x_m - x_n\| &\leq \|x_m - x_{m-1}\| + \|x_{m-1} - x_{m-2}\| + \cdots + \|x_{n+1} - x_n\| \\
                              &\leq \gamma^{m-1}\|x_1-x_0\| + \gamma^{m-2}\|x_1-x_0\| + \dots + \gamma^{n}\|x_1-x_0\| \\
                              &= \|x_1-x_0\| (\gamma^{m-1} + \gamma^{m-2} + \dots + \gamma^n) \\
                              &= \|x_1-x_0\| \frac{\gamma^n(1-\gamma^{m-n})}{1-\gamma} \\
                              &\leq \|x_1-x_0\| \frac{\gamma^n}{1-\gamma}
            \end{align}
            where $m \geq n$, without loss of generality. Since $\gamma \in (0,1)$, the last line converges to 0. That is, the sequence $\{x_k\}$ is a Cauchy sequence.

            \textit{Part 2.} (Existence) $\lim_{k \to \infty} x_k$ is a solution of fixed-point equation $x=f(x)$.

            Let $x^* = \lim_{k \to \infty} x_k$. Then, following holds.
            \begin{align}
                \|x^*-f(x^*)\| &\leq \|x^*-x_k\| + \|f(x_k)-f(x^*)\| + \|x_k-f(x_k)\| \\
                               &\leq \|x^*-x_k\| + \gamma \|x_k - x^*\| + \|x_k - f(x_k)\|
            \end{align}
            Using the results of \textit{Part 1} and squeeze theorem, we can show that $x^*=f(x^*)$.

            \textit{Part 3.} (Uniqueness) The fixed-point $x^*$ is unique.

            Suppose there is another fixed point $x'$. Then, $\|x^*-x'\| = \|f(x^*)-f(x')\| \leq \gamma \|x^*-x'\|$. Since $\gamma \in (0,1)$, the only possibility is $x^*=x'$.


        \end{proof}
    \subsection{Contraction Property of Bellman Optimality Equation}
        Now, to use the contraction mapping theorem, we prove that the right-hand side of BOE \ref{eq:BOERightHandSide} is a contraction mapping.
        \begin{thm}{Contraction Property of Bellman Optimality Equation}{ContPropBellmanOptEq}
            Let the function $f$ be that of equation \ref{eq:BOERightHandSide}. Then, the following contraction mapping property holds for any $v_1,v_2 \in \mathbb{R}^{|\mathcal S|}$:
            \begin{equation}
                \|f(v_1)-f(v_2)\|_{\infty} \leq \gamma \|v_1 - v_2\|_\infty
            \end{equation}
            where $\gamma \in (0,1)$ is the discount rate, and $\|\|_\infty$ is the maximum norm, which is the maximum absolute value of the elements in a vector.
        \end{thm}
        \begin{proof}
            Let $v_1, v_2 \in \mathbb{R}^{\mathcal{S}}$, and let $\pi_1^* = \arg\max_\pi{(r_\pi + \gamma P_\pi v_1)}$, $\pi_2^* = \arg\max_\pi{(r_\pi + \gamma P_\pi v_2)}$.
            Then,
            \begin{align}
                f(v_1) &= \max_\pi{(r_\pi + \gamma P_\pi v_1)} = r_{\pi_1^*} + \gamma P_{\pi_1^*}v_1 \geq r_{\pi_2^*} + \gamma P_{\pi_2^*}v_1 \\
                f(v_2) &= \max_\pi{(r_\pi + \gamma P_\pi v_2)} = r_{\pi_2^*} + \gamma P_{\pi_2^*}v_2 \geq r_{\pi_1^*} + \gamma P_{\pi_1^*}v_2
            \end{align}
            where $\geq$ is applied elementwise. This result leads to
            \begin{align}
                f(v_1)-f(v_2) &\leq \gamma P_{\pi_1^*}(v_1-v_2) \\
                f(v_2)-f(v_1) &\leq \gamma P_{\pi_2^*}(v_2-v_1)
            \end{align}
            Let $z=\max{\{|\gamma P_{\pi_1^*}(v_1-v_2)|, |\gamma P_{\pi_2^*}(v_2-v_1)|\}}$, where $\max(\cdot)$ and $|\cdot|$ are all applied elementwise.
            Then,
            \begin{equation}
                \|f(v_1) - f(v_2)\|_{\infty} \leq \|z\|_{\infty}. \label{tmpeq:1}
            \end{equation}
            Now, let $z_i$ be the $i$-th entry of $z$, and let $p_i^{\top}$ and $q_i^{\top}$ be the $i$-th row of $P_{\pi_1^*}$ and $P_{\pi_2^*}$, respectively.
            Then, $z_i = \max{\{\gamma |p_i^{\top}(v_1-v_2)|, \gamma|q_i^{\top}(v_1-v_2)|\}}$.
            Since all the elements of vector $p_i$ is nonnegative and thier sum equals 1,
            \begin{equation}
                |p_i^\top(v_1-v_2)| \leq p_i^\top|v_1-v_2| \leq \|v_1-v_2\|_\infty.
            \end{equation}
            where $|\cdot|$ is applied elementwise.
            Similarly, $|q_i^\top(v_1-v_2)| \leq \|v_1-v_2\|_\infty$. Therefore $\|z\|_\infty \leq \gamma \|v_1-v_2\|_\infty$.
            Substituting to \ref{tmpeq:1},
            \begin{equation}
                \|f(v_1)-f(v_2)\|_{\infty} \leq \gamma \|v_1 - v_2\|_\infty.
            \end{equation}
        \end{proof}
\newpage
\section{Optimal Policy, Optimal State Value and Bellman Optimality Equation}\label{sec:OptPolOptStateValBOE}
    We have a following Theorem of BOE:
    \begin{thm}{Solution of the Bellman Optimality Equation}{SolutionBellmanOptEq}
        Let $f$ be as the equation \ref{eq:BOERightHandSide}. Then, the BOE $v=f(v)$ has following properties:
        \begin{enumerate}
            \item There exists a unique solution;
            \item The solution can be obtained iteratively by $v_{k+1}=f(v_k)$.
        \end{enumerate}
    \end{thm}
    \begin{proof}
        All the previous Theorems (\ref{thm:ContractMap}, \ref{thm:ContPropBellmanOptEq}) proves this trivially.
    \end{proof}

    Now, we should prove that the solution of the Bellman optimality equation is really an optimal state-value function,
    and prove that the greedy policy derived from it is the optimal policy.
    \begin{thm}{Optimality of $v^*$ and $\pi^*$}{OptPolOptStateVal}
        Let $v^*$ be the solution of BOE $v=f(v)$, and let $\pi^* = \arg\max_{\pi \in \Pi}(r_\pi + \gamma P_\pi v^*)$.
        The, the $v^*$ is an optimal state vaule function, and the $\pi^*$ is an optimal policy. That is, for any policy $\pi$,
        following holds;
        \begin{equation}
            v^* = v_{\pi^*} \geq v_\pi
        \end{equation}
        where $v_\pi$ is the state value function of $\pi$, and $\geq$ is applied elementwise.
    \end{thm}
    \begin{proof}
        For any policy $\pi$, $v_\pi = r_\pi + \gamma P_\pi v_\pi$ holds (Bellman equation). Then,
        \begin{align}
            v^* &= \max_{\tau \in \Pi}(r_\tau + \gamma P_\tau v^*) \\
                &= r_{\pi^*} + \gamma P_{\pi^*} v^* \\
                &\geq r_\pi + \gamma P_\pi v^*
        \end{align}
        and we have
        \begin{equation}
            v^*-v_\pi \geq \gamma P_\pi(v^*-v_\pi).
        \end{equation}
        Using this inequality repeatedly, we have $v^*-v_\pi \geq \gamma^n P_\pi^n (v^*-v_\pi)$ for any $n \in \mathbb{N}$.
        Since $\gamma \in (0,1)$ and $P_\pi$ is a nonnegative matrix with all its elements less than or equal to 1,
        $\lim_{n \to \infty} \gamma^n P_\pi^n (v^*-v_\pi)=0$. That is, $v^*-v_\pi \geq \lim_{n \to \infty} \gamma^n P_\pi^n (v^*-v_\pi)=0$.
    \end{proof}
    The following theorem examines the optimal policy more precisely.
    \begin{thm}{Greedy Optimal Policy}{GreedyOptPol}
        For any $s \in \mathcal{S}$, the following deterministic policy is an optimal policy for solving the BOE.
        \begin{equation}
            \pi^*(a|s) = 
            \begin{cases}
                1, &a = \arg\max_a q^*(a,s) \\
                0, &a \neq \arg\max_a q^*(a,s)
            \end{cases}
        \end{equation}
        where $q^*(s,a) = \sum_{r \in \mathcal{R}}p(r|s,a)r + \gamma \sum_{s' \in \mathcal{S}}p(s'|s,a)v^*(s')$.
    \end{thm}
    \begin{proof}
        From Theorem~\ref{thm:SolutionBellmanOptEq} and~\ref{defn:BellOptEq}, we see that
        \begin{align}
            v^*(s) &= \max_{\pi \in \Pi}\sum_{a \in \mathcal{A}}\pi(a|s)q^*(s,a) \\
                        &\leq \left(\max_{\pi \in \Pi}\sum_{a \in \mathcal{A}}\pi(a|s)\right) \max_{a'} q^*(s,a') \label{tmpeq:2} \\
                        &= \max_{a'}q^*(s,a')
        \end{align}
        and the given deterministic policy $\pi^*$ satisfies the equality of inequality~\ref{tmpeq:2}.
    \end{proof}
    The important truth is that the optimal state value is unique but the optimal policy is not. The Theorem~\ref{thm:GreedyOptPol}
    only shows that there always exists at least one deterministic optimal policy. There still can be other optimal policies, too, stochastic or whatever.

\section{Factors that Influence Optimal Policies}
    Now, we analyze the factors of BOE that influences optimal policies. The factors are: reward and discount rate.
    \subsection{Reward}
    \subsection{Discount Rate}